---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r LoadData, echo = TRUE, cache = TRUE}
library(data.table)
library(lubridate)

#read data, and 
activity <- read.csv("activity.csv", header = TRUE)
activity_no_na = na.omit(activity)
activity_no_na <- data.table(activity_no_na)
activity_no_na$date <- ymd(activity_no_na$date)
```

## What is mean total number of steps taken per day?

```{r Histogram, echo = TRUE}
# group by date and sum the steps
agg <- aggregate(steps ~ date, activity_no_na, sum)
hist(agg$steps, main = "Total Steps per Day", col = "brown", xlab = "Steps")
```

# Mean and Median
```{r CalculateMean, echo = TRUE}
meanSteps <- mean(agg$steps, na.rm = TRUE)
medSteps <- median(agg$steps, na.rm = TRUE)
```
The mean of the total number of steps is **`r toString(meanSteps)`**.  
The median of the total number of steps is **`r medSteps`**.


## What is the average daily activity pattern?
```{r AvgSteps, echo = TRUE}
avgSteps <- aggregate(steps ~ interval, activity_no_na, mean)
plot(avgSteps$interval, avgSteps$steps, type = "l", xlab = "Interval", ylab = "Average Steps", 
     main = "Average Daily Activity", col = "brown")

# max # of avg steps
maxSteps <- max(avgSteps$steps)
maxSteps <- format(round(maxSteps, 2), nsmall = 2)
maxInterval <- subset(avgSteps, steps == max(steps))$interval
```
The interval **`r toString(maxInterval)`** contains the maximum number of steps **`r maxSteps`**.

## Imputing missing values
```{r NAs, echo = TRUE}

# calculate number of NAs
naRows <- subset(activity, is.na(activity))
completeRows <- subset(activity, !is.na(activity))

naCount <- nrow(naRows)

merged <- merge(naRows, avgSteps, by = "interval")
naRows$steps <- merged$steps.y
final <- rbind(naRows, completeRows)
finalAgg <- aggregate(steps ~ date, final, sum)
hist(finalAgg$steps, main = "Total Steps per Day w/ Imputed Values", col = "red", xlab = "Steps")

# mean and median
meanFinalSteps <- mean(finalAgg$steps, na.rm = TRUE)
meanFinalSteps <- format(round(meanFinalSteps, 2), nsmall = 2)
medFinalSteps <- median(finalAgg$steps, na.rm = TRUE)
medFinalSteps <- format(round(medFinalSteps, 2), nsmall = 2)
```

There are **`r toString(naCount)`** missing values.  
The mean for the dataset with imputed values is **`r meanFinalSteps`**.  
The median for the dataset with imputed values is **`r medFinalSteps`**.  
The mean of the complete is the same as the mean for the dataset without missing values. The median is almost the same, with a difference of 250.

## Are there differences in activity patterns between weekdays and weekends?
