---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r LoadData, echo = TRUE, cache = TRUE}
library(data.table)
library(lubridate)

#read data, and 
activity <- read.csv("activity.csv", header = TRUE)
activity = na.omit(activity)
activity <- data.table(activity)
activity$date <- ymd(activity$date)
```

## What is mean total number of steps taken per day?

```{r Histogram, echo = TRUE}
# group by date and sum the steps
agg <- aggregate(steps ~ date, activity, sum)
hist(agg$steps, main = "Total Steps per Day", col = "brown", xlab = "Steps")
```

# Mean and Median
```{r CalculateMean, echo = TRUE}
meanSteps <- mean(agg$steps, na.rm = TRUE)
medSteps <- median(agg$steps, na.rm = TRUE)
```
The mean of the total number of steps is **`r toString(meanSteps)`**.  
The median of the total number of steps is **`r medSteps`**.


## What is the average daily activity pattern?
```{r AvgSteps, echo = TRUE}
avgSteps <- aggregate(steps ~ interval, activity, mean)
plot(avgSteps$interval, avgSteps$steps, type = "l", xlab = "Interval", ylab = "Average Steps", 
     main = "Average Daily Activity", col = "brown")

# max # of avg steps
maxSteps <- max(avgSteps$steps)
maxSteps <- format(round(maxSteps, 2), nsmall = 2)
maxInterval <- subset(avgSteps, steps == max(steps))$interval
```
The interval **`r toString(maxInterval)`** contains the maximum number of steps **`r maxSteps`**.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
